### 2016-09-14 introduction.tex
Corrections up to page 3

The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on a global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, and smart grids. The number of wireless devices is expected to increase by five times up to 50 billion devices \cite{cisco2011forecast}. It is generally believed, that those devices will not be connected in the same manner as current devices are connected today. Centralized system will collapse in terms of capacity, while distributed systems appear as an alternative. Therefore the communication architecture in future systems will be dominated by mesh technologies. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high \ac{QoE} for the final user. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area. This not only allows to seamlessly develop implementations that are compatible with higher end devices, but also permits the reuse of stable software from widely-used \ac{OS}. For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications. Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance. However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading. Moreover, this work does not consider possible network coding applications. Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of a public road isle of lamps and reports it to a monitoring center. A use case regarding remote environment surveillance is presented in \cite{leccese2014newimaging}, where the \ac{Raspi} reports air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves. This data is sent to a monitoring center to ensure the preservation of the paintings. Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints. Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}. Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users. The current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.

>>>>

The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on a global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, and smart grids. The number of wireless devices is expected to increase by five times up to 50 billion devices \cite{cisco2011forecast}. It is generally believed, that those devices will not be connected in the same manner as current devices are connected today. Centralized systems will collapse in terms of capacity, while distributed systems appear as an alternative. Therefore, we believe mesh technologies will play a major role in the communication architecture in future systems. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high \ac{QoE} for the final user. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

The emergence of powerful and inexpensive single-board computers opens new possibilities in this area. By running a standard \ac{OS}, they allow implementations that are compatible with higher end devices. Also, they utilize stable software supported by their communities. For example, the Iridis-pi platform \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications. Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance. This work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading. Moreover, this work does not consider possible network coding applications. Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of an isle of lamps on a public road and reports it to a monitoring center. A use case regarding remote environment surveillance using the Raspi and the Arduino \cite{arduino} tehnologies is presented in \cite{leccese2014newimaging}. Here both devices report air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves. This data is sent to a monitoring center to ensure the preservation of the paintings. Further, authors in \cite{sapes2016finger} consider FingerScanner, a technology that utilizes the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints. Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}. These applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users. The current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.

### 2016-09-14 overview.tex
Corrections up to page 3

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a 8 GB \ac{SD} memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and academic staff
to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. The
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s are configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP} address. This
eliminates the demand for monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

@@

We will refer to the
testbed administrator as the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedures are
performed by the testbed administrator in a regular \ac{PC} running a
Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to store it later in a memory card for each \ac{Raspi}.
Once configured, we store the resulting image file in a \ac{HTTP} server
as backup and in case the testbed administrator requires to make new
changes to this file. We also put all the required configuration files
and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is
a single place where system setup is stored and could be modified.
This simplifies the system maintenance.
As
%another design consideration, when working on a testbed with many devices,
it may not always be desirable to make persistent changes
on the \ac{Raspi}s. For example when different users are interested in
running experiments on a rebooted testbed. We later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.

>>>>

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a 8 GB \ac{SD} memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and academic staff
to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. The
testbed description and procedures for setting it up are not restricted
to this academic scenario. All \ac{Raspi}s are configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP} address. This
eliminates the demand for monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

@@

We will refer to the
testbed administrator as the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedures are
performed by the testbed administrator in a \ac{PC} running a
Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to store it later on a memory card for each \ac{Raspi}.
Once configured, we store the resulting image file in a \ac{HTTP} server
as backup and in case the testbed administrator requires to make new
changes to this file. For our case, we use the university \ac{HTTP} server but any other servers can be used. We also put all the required configuration files
and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is
a single place where system setup is stored and could be modified.
This simplifies the system maintenance.
As
%another design consideration, when working on a testbed with many devices,
it may not always be desirable to make persistent changes
on the \ac{Raspi}s. For example when different users are interested in
running experiments on a rebooted testbed. We later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.
